---
title: "Lahman SQL/R Exploration"
output: html_notebook
---

```{r, echo = FALSE, warning = FALSE}
library(tidyverse)
library(DBI)
library(RPostgres)
library(RcppRoll)

con <- dbConnect(Postgres(),                 
                 dbname = 'lahman',
                 host = 'localhost',
                 port = 5432, 
                 user = 'postgres',
                 password = rstudioapi::askForPassword("Database Password: "))
```

1. Is there any correlation between number of wins and team salary? Use data from 2000 and later to answer this question. As you do this analysis, keep in mind that salaries across the whole league tend to increase together, so you may want to look on a year-by-year basis. **There is a moderately positive correlation between team salaries and wins by year, though this was lower between 2008 and 2015, with a jump up in 2016.**

```{r, echo = FALSE, warning = FALSE}
query_question1 = "
WITH total_salaries AS (
	SELECT
		yearid,
		teamid,
		SUM(salary) AS total_team_salary_that_year
	FROM salaries
	GROUP BY yearid, teamid
),
team_wins AS (
	SELECT
		yearid,
		teamid,
		w,
		name
	FROM teams
	WHERE yearid >= 2000
)

SELECT
	ts.yearid,
	total_team_salary_that_year,
	w,
	tw.name
FROM total_salaries AS ts
INNER JOIN team_wins AS tw
ON ts.yearid = tw.yearid
AND ts.teamid = tw.teamid
ORDER BY ts.yearid;
"

res_question1 <- dbSendQuery(con, query_question1)
salaries_and_wins <- dbFetch(res_question1)

salaries_and_wins %>% 
  group_by(yearid) %>% 
  mutate(correlation_by_year = cor(x = total_team_salary_that_year, y = w),
         .keep = "none") %>%
  distinct() %>% 
  ggplot(aes(yearid, correlation_by_year)) +
  geom_line(color = 'red') +
  geom_point() +
  labs(x = "Year",
       y = "Correlation Value",
       title = "Correlation of Team Salary and Wins Over Time") +
  ylim(0,1) +
  scale_x_continuous(limits = c(2000,2016),
                     breaks = seq(2000,2016,1))
```

2. In this question, you will explore the connection between number of wins and attendance. 

a. Does there appear to be any correlation between attendance at home games and number of wins? **There is a moderately positive correlation between home game attendance and wins by year. This was strongest in the 50's and 60's and bottomed out in the early 70's but has since returned. This is probably best interpreted as meaning that winning teams enjoy more fans at home games, rather than the other way around.**

```{r, echo = FALSE, warning = FALSE}
query_question2a = "
WITH team_wins AS (
	SELECT
		yearid,
		teamid,
		w,
		name
	FROM teams
),
home_game_attendance AS (
SELECT 
	year,
	team,
	ROUND(SUM(attendance)/SUM(games)::numeric,2) AS attendance_per_home_game
FROM homegames
GROUP BY year, team
HAVING SUM(attendance)/SUM(games)::numeric > 0
ORDER BY year
)
SELECT
	year,
	name,
	w,
	attendance_per_home_game
FROM team_wins AS tw
INNER JOIN home_game_attendance AS hga
ON tw.yearid = hga.year
AND tw.teamid = hga.team;
"
res_question2a <- dbSendQuery(con, query_question2a)
wins_and_attendance <- dbFetch(res_question2a) %>% as_tibble()

wanda <- wins_and_attendance %>% 
  group_by(year) %>% 
  mutate(correlation_by_year = cor(x = attendance_per_home_game, y = w), 
         .keep = "none") %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(correlation_by_year = roll_mean(correlation_by_year,
                                        fill = NA,
                                        n = 10L, 
                                        align = 'right',
                                        na.rm = T)) %>% 
  slice(10:n())

wanda %>% 
  ggplot(aes(year, correlation_by_year)) +
  geom_line(color = 'red') +
  labs(x = "Year",
       y = "Mean Correlation as of Last 10 Years",
       title = "Correlation of Home Game Attendance and Wins Over Time For All Teams") + 
  ylim(0,1) +
  scale_x_continuous(limits = c(1890,2020),
                     breaks = seq(1890,2020,10))


```

b1. Do teams that win the world series see a boost in attendance the following year? **Generally, no. The distribution of attendance boost or reduction numbers is normal with a median at zero.**

```{r, echo = FALSE, warning = FALSE}

query_question2b1 = "
WITH team_ws_win AS (
	SELECT
		yearid,
		teamid,
		CASE WHEN wswin = 'Y' THEN 'World Series Winner'
		ELSE 'Not World Series Winner'
		END AS ws_win,
		name
	FROM teams
),
home_game_attendance AS (
	SELECT 
		year,
		team,
		LEAD(ROUND(SUM(attendance)/SUM(games)::numeric,2), 1) 
		OVER(PARTITION BY team ORDER BY year) - ROUND(SUM(attendance)/SUM(games)::numeric,2) AS attendance_boost
	FROM homegames
	GROUP BY year, team
	HAVING SUM(attendance)/SUM(games)::numeric > 0
	ORDER BY team
)
SELECT
	year,
	name,
	ws_win,
	attendance_boost
FROM team_ws_win AS tw
INNER JOIN home_game_attendance AS hga
ON tw.yearid = hga.year
AND tw.teamid = hga.team
WHERE ws_win = 'World Series Winner'
AND year <> 2016
ORDER BY year;
"

res_question2b1 <- dbSendQuery(con, query_question2b1)
ws_and_attendance <- dbFetch(res_question2b1)

robust_scalar_function <- function(variable_in_question){
  return (variable_in_question - median(variable_in_question, na.rm = T)) /(quantile(variable_in_question, probs = .75) - quantile(variable_in_question, probs = .25))
}

ws_and_attendance %>% 
  mutate(attendance_boost_scaled = robust_scalar_function(attendance_boost)) %>% 
  ggplot(aes(attendance_boost_scaled)) +
  geom_density(color = "black", fill = "darkgreen", alpha = 0.2) +
  labs(x = "Attendance Boost Numbers",
       y = "Density",
       title = "Density of Attendance Boost Distribution Given World Series Win") + 
  scale_x_continuous(limits = c(-15000,15000),
                     breaks = seq(-15000,15000,5000))

```

b2. What about teams that made the playoffs? Making the playoffs means either being a division winner or a wild card winner. **Same as the answer above: there's effectively no correlation.**


```{r, echo = FALSE, warning = FALSE}

query_question2b2 = "
WITH team_playoffs AS (
	SELECT
		yearid,
		teamid,
		CASE WHEN divwin = 'Y' OR wcwin = 'Y' THEN 'Playoffs'
		ELSE 'No Playoffs'
		END AS playoffs,
		name
	FROM teams
),
home_game_attendance AS (
	SELECT 
		year,
		team,
		LEAD(ROUND(SUM(attendance)/SUM(games)::numeric,2), 1) 
		OVER(PARTITION BY team ORDER BY year) - ROUND(SUM(attendance)/SUM(games)::numeric,2) AS attendance_boost
	FROM homegames
	GROUP BY year, team
	HAVING SUM(attendance)/SUM(games)::numeric > 0
	ORDER BY team
)
SELECT
	year,
	name,
	playoffs,
	attendance_boost
FROM team_playoffs AS tp
INNER JOIN home_game_attendance AS hga
ON tp.yearid = hga.year
AND tp.teamid = hga.team
WHERE playoffs = 'Playoffs'
AND year <> 2016
ORDER BY year;
"
res_question2b2 <- dbSendQuery(con, query_question2b2)
playoffs_and_attendance <- dbFetch(res_question2b2)

playoffs_and_attendance %>% 
  mutate(attendance_boost_scaled = robust_scalar_function(attendance_boost)) %>% 
  ggplot(aes(attendance_boost_scaled)) +
  geom_density(color = "black", fill = "darkgreen", alpha = 0.2) +
  labs(x = "Attendance Boost Numbers",
       y = "Density",
       title = "Density of Attendance Boost Distribution Given Playoffs Participation") + 
  scale_x_continuous(limits = c(-15000,15000),
                     breaks = seq(-15000,15000,5000))

```


